---
title: "Beer Case Study"
author: "Data Busters, Inc. (Griffin J. Lane, Kristopher Henley, Josephine McDaniel)"
date: "October 18, 2017"
output: html_document
---

#Beer Case Study
###The client, Mega Awesome Beer Corp. (MABco), wants us to...  etc etc... 

# 1a. import a csv file include the column name

```{r, echo=TRUE}
BeerData <- read.csv2(file="C:/Users/Griff/OneDrive/Documents/!_SKOOL/Doing Data Science/Teamwork/Case-Study-1/MSDS Project/Beers.csv", header=TRUE, sep=",", col.name=c("Name", "BeerID", "ABV", "IBU", "BreweryID", "Style", "Ounces"))

BrewiesData <- read.csv2(file="C:/Users/Griff/OneDrive/Documents/!_SKOOL/Doing Data Science/Teamwork/Case-Study-1/MSDS Project/Breweries.csv", header=TRUE, sep=",", col.name=c("BreweryID", "BreweryName", "City", "State"))
```

# 1b. display the summary and structure of the file

```{r}
summary (BeerData)
str(BeerData)
summary (BrewiesData)
str(BrewiesData)
```

# merge the files
```{r,  echo=TRUE}
merged <- merge(BeerData, BrewiesData, by = "BreweryID", all=FALSE)
```


#  top and bottom 6 of merge file
```{r, echo=TRUE}
head(merged)
tail(merged)
```

# display number of NAs

```{r, echo=TRUE}
##First we must fix the numeric columns incorrectly listed as ordered factors
##A direct conversion to a numeric will instead return the rank of the factor
##Converting to a matrix first avoids this issue
merged$ABV <- as.numeric(as.matrix(merged$ABV))
merged$Ounces <- as.numeric(merged$Ounces)



colSums(is.na(merged))
```

# sort by city

```{r, echo=TRUE}
beer <- merged[order(merged$City),] 
```

# How many breweries are present in each state
```{r, echo=TRUE}
table(beer$State[beer$BreweryID])
```


###The client also wishes to examine the median alcohol content and IBUs for each state
```{r}

##Create columns for both measurements to allow for quick lookups in the future on a per tuple basis
beer$StateABVMedian <- sapply(beer$ABV[beer$State], median)
beer$StateIBUMedian <- sapply(beer$IBU[beer$State], median)

##Create a list of every state included and store it in a variable
allStates <- unique(beer$State[beer$BreweryID])

##Use that list to index the new columns and create a data frame to display the findings
mediansByState <- data.frame(State=allStates, ABV_median = beer$StateABVMedian[allStates], IBU_median = beer$StateIBUMedian[allStates])

##Display the ne data frame that holds the median data
mediansByState
``` 

###Visualization
```{r}
library(ggplot2)
##Create a plot with the staes along the x-axis
p <- ggplot2::ggplot(mediansByState, aes(x = State))

##Create and add the y-axis. Also, give it a low alpha value and change it's fill color
p <- p + geom_bar(aes(y = IBU_median, color = "IBU"), stat="identity", alpha = .5 , fill="blue")

##Create and add the second y-axis. Also, give it a low alpha value and change it's fill color
##Note that the extra 100 multiplier is to turn the scale of those data into a percentage scale
p <- p + geom_bar(aes(y = ABV_median*100*6, color = "ABV"), alpha = .5 , fill="red", stat="identity")

##Create the right side second axis and give it the inverse scale used to transform the data
p <- p + scale_y_continuous(sec.axis = sec_axis(~./6, name = "ABV")) 

##Display the ggplot and marvel at its splendor
p
```



###For their records, the client has also asked that we determine which state sells the most bitter beer as well as the most alcoholic beer
```{r}
beer$State[which.max(beer$ABV)]

beer$State[which.max(beer$IBU)]
```

