---
title: "Mega Awesome Beer Corp. (MABco) Case Study"
author: "Data Busters Inc. (K Henley, G Lane, J McDaniel)"
date: "October 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Mega Awesome Beer Corp. (MABco), has commissioned Data Busters, Inc.(DBI) to examine data sets they have provided, with the purpose of answering some key questions and findings that should be derived directly from the data provided.  These are the results. 

Two data sets were provided by MABco and used by DBI to provide the findings below.  One data 
set provides background on the beer offerings currently in production at MABco, and the other
provides brewery data where each beer is produced.  We've read the raw data into our systems
for analysis.

```{r, echo=TRUE}
BeerData <- read.csv2(file="C:/Users/henlek1/Desktop/Beers.csv", header=TRUE, sep=",", col.name=c("Name", "BeerID", "ABV", "IBU", "BreweryID", "Style", "Ounces"))

BrewiesData <- read.csv2(file="C:/Users/henlek1/Desktop/Breweries.csv", header=TRUE, sep=",", col.name=c("BreweryID", "BreweryName", "City", "State"))
```

Below is the summary for both data sets.  As you can see the data provided is organized
into 2410 types of beer, and between the two sets of data we have 11 measurable variables 
to work with.  We will walk you through the results below, and provide you with the
background necessary to understand the context behind our findings, and hopefully answer
all of your questions.
```{r, echo=TRUE}
summary (BeerData)
str(BeerData)
summary (BrewiesData)
str(BrewiesData)
```

In order to take full advantage of the data provided we have opted to merge the two files.
The new merged data set contains ALL the measurable variables in one file called "merged",
but will now allow us to cross reference the beer and brewery data for a more complete analysis.

```{r,  echo=TRUE}
merged <- merge(BeerData, BrewiesData, by = "BreweryID", all=FALSE)
```

To further organize our data we have taken a look at the first and last few entries within
our newly merged data set.  This will allow us to a) ensure we have a better understanding of
the new data file we are working with, and b) help to ensure all data was properly pulled 
into the file as expected.  Upon inspection of the first 6 and last 6 entries within the 'merged' file we see no unexpected issues.

```{r, echo=TRUE}
head(merged)
tail(merged)
```

As expected, we did encounter some instances of missing data after the merge.  Specifically, there were 1067 instances in total of missing data that were being incorrectly measured as "NA".  There were 62 instances in the ABV column and 1005 in the IBU column.  

```{r, echo=TRUE}
##First we must fix the numeric columns incorrectly listed as ordered factors
##A direct conversion to a numeric will instead return the rank of the factor
##Converting to a matrix first avoids this issue
merged$ABV <- as.numeric(as.matrix(merged$ABV))
merged$Ounces <- as.numeric(merged$Ounces)



colSums(is.na(merged))
```

Now that we've organized the data provided we can get down to providing the information that
you've requested.  In order to further understand MABco's brewery presence in each U.S. state we have provided a summary table below:

```{r, echo=TRUE}
beer <- merged[order(merged$City),]
table(beer$State[beer$BreweryID])
```

DBI was also tasked with providing a better understanding of MABco's beer offerings in
terms of alcohol content (ABV).  We determined that the highest median ABV is 0.092, and
was reported in the states of Georgia, Florida, and Idaho when computing the medians for all beer offerings in each state respectively.

We have also provided median IBU values for each state for reference in the table below.
```{r, echo=TRUE}

##Create columns for both measurements to allow for quick lookups in the future on a per tuple basis
beer$StateABVMedian <- sapply(beer$ABV[beer$State], median)
beer$StateIBUMedian <- sapply(beer$IBU[beer$State], median)

##Create a list of every state included and store it in a variable
allStates <- unique(beer$State[beer$BreweryID])

##Use that list to index the new columns and create a data frame to display the findings
mediansByState <- data.frame(State=allStates, ABV_median = beer$StateABVMedian[allStates], IBU_median = beer$StateIBUMedian[allStates])

##Display the ne data frame that holds the median data
mediansByState
``` 

To help MABco better understand these results we've also plotted them graphically for your reference.
```{r, echo=TRUE}
library(ggplot2)
##Create a plot with the staes along the x-axis
p <- ggplot2::ggplot(mediansByState, aes(x = State))

##Create and add the y-axis. Also, give it a low alpha value and change it's fill color
p <- p + geom_bar(aes(y = IBU_median, color = "IBU"), stat="identity", alpha = .5 , fill="blue")

##Create and add the second y-axis. Also, give it a low alpha value and change it's fill color
##Note that the extra 100 multiplier is to turn the scale of those data into a percentage scale
p <- p + geom_bar(aes(y = ABV_median*100*6, color = "ABV"), alpha = .5 , fill="red", stat="identity")

##Create the right side second axis and give it the inverse scale used to transform the data
p <- p + scale_y_continuous(sec.axis = sec_axis(~./6, name = "ABV")) 

##Display the ggplot and marvel at its splendor
p
```


Upon review of the results we derived from looking at the median alcohol content (ABV),
as well as the median bitterness levels (IBU) we are able to confirm that the states with
beer offerings with the highest median ABV are Georgia, Florida, and Idaho (with median ABVs of 0.092 respectively) as previously confirmed.  

**However the beer with the highest ABV is Lee Hill Series Vol. 5 - 
Belgian Style Quadrupel Ale (0.128 ABV).  This beer is sold in COLORADO.**
```{r, echo=TRUE}
beer$State[which.max(beer$ABV)]
```

We are further able to confirm that the states with the highest median bitter beers as measured in IBU's are NY, OH, NC, OR, PA, NM, MT, SC, and NJ respectively, with each state returning
a median IBU score of 135.  
**However, the individual state returning the highest IBU (138)is OREGON (Bitter Bitch Imperial IPA)**

```{r, echo=TRUE}
beer$State[which.max(beer$IBU)]
```

```{r, echo=TRUE}
summary(beer$ABV)
```



*What is the relationship of Alcohol content and Bitterness for MABco's beer offerings?*
The graph below is a scatterplot that examines the relationship between ABV and IBU for each
beer that MABco has provided us data on.  As you can see there is a direct linear correlation 
betwen the bitterness of the beer and it's corresponding alcohol content.  As the alcohol content increases, so does the bitterness.


```{r, echo=TRUE}
## Scatterplot examining relationship between bitterness and alcohol content ##
## ABV vs. IBU 
library(car)
scatterplot(ABV ~ IBU, data=merged, 
  	xlab="Alcohol Content in ABV", ylab="Bitterness in IBU", 
   main="Relationship between ABV and IBU", 
   labels=row.names(merged))
```



In summary DBI has provided MABco with a look at it's brewery presence (by U.S. State), and the results show the top five to be Texas (255), Colorado (253), Indiana (199), Oregon (188), and New York (168).  We've also provided data regarding the median Alcohol by volume (ABV) by each state.  Our findings determined that the states with the highest median (ABV) are Georgia, Florida, and Idaho, each recording a median of 0.092 respectively.  We also called out that although these states recorded the highest median values, the state with highest overall ABV from a single beer offering was found outside of this population.  Colorado sells the beer with the overall highest ABV offered (Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale (0.128 ABV)).  When examining all beer offerings the overall average (ABV) was found to be 0.05977.  Further, we provided data regarding the bitterness of MABco's beer offerings as measured in (IBU), and reported our findings by state. DBI determined that the states with the highest median bitter beers as measured in IBU's are NY, OH, NC, OR, PA, NM, MT, SC, and NJ respectively, with each state returning a median IBU score of (135 IBU).  When measuring (IBU) overall maximum's we found that the individual state returning the highest IBU (138) was Oregon (Bitter Bitch Imperial IPA).  Please note the state with the maximum bitterness value does reside in the highest median population we examined (unlike the ABV findings).  Lastly, we provided a graphical representation outlining the relationship between the Alcohol by Volume (ABV) and the bitterness (IBU) of each of MABco's beer offerings.  The graph clearly outlined a positive linear relationship between the variables.  In general, as the ABV increased, so did the bitterness levels as recorded by IBU.  Further evaluation can be conducted here with MABco's direction if more detailed analysis is needed regarding the noted correlation.

This concludes our analysis.
