---
title: "Mega Awesome Beer Corp. (MABco) Case Study"
author: "Data Busters Inc. (K Henley, G Lane, J McDaniel)"
date: "October 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Mega Awesome Beer Corp. (MABco), has commissioned Data Busters, Inc.(DBI) to examine data sets they have provided, with the purpose of answering some key questions and findings that should be derived directly from the data provided.  These are the results. 

# Two data sets were provided by MABco and used by DBI to provide the findings below.

```{r, echo=TRUE}
BeerData <- read.csv2(file="C:/Users/henlek1/Desktop/Beers.csv", header=TRUE, sep=",", col.name=c("Name", "BeerID", "ABV", "IBU", "BreweryID", "Style", "Ounces"))

BrewiesData <- read.csv2(file="C:/Users/henlek1/Desktop/Breweries.csv", header=TRUE, sep=",", col.name=c("BreweryID", "BreweryName", "City", "State"))
```

# Below is the summary for both data sets.  As you can see the data provided is organized
# into 2410 types of beer, and between the two sets of data we have 11 measurable variables 
# to work with.  We will walk you through the results below, and provide you with the
# background necessary to understand the context behind our findings.
```{r, echo=TRUE}
summary (BeerData)
str(BeerData)
summary (BrewiesData)
str(BrewiesData)
```

# In order to take full advantage of the data provided we have opted to merge the two files.
# The new merged data set contains ALL the measurable variables in one file called "merged"

```{r,  echo=TRUE}
merged <- merge(BeerData, BrewiesData, by = "BreweryID", all=FALSE)
```

# To further organize our data we have taken a look at the first and last few entries within
# our newly merged data set.  This will allow us to a) ensure we have a better understanding of
# the new data file we are working with, and b) help to ensure all data was properly pulled 
# into the file as expected.  Upon inspection of the 'merge' file we see no unexpected issues.

```{r, echo=TRUE}
head(merged)
tail(merged)
```

## I think we need to provide an NA total here prior to correcting them? ##
# As expected, we do encounter some missing data that are being incorrectly measured as "NA"
# This is not unusual, and we are able to reformat the data with some simple coding behind the
# scenes that reassigns anything with an unmeasurable "NA" to a zero (0) as appropriate.

```{r, echo=TRUE}
##First we must fix the numeric columns incorrectly listed as ordered factors
##A direct conversion to a numeric will instead return the rank of the factor
##Converting to a matrix first avoids this issue
merged$ABV <- as.numeric(as.matrix(merged$ABV))
merged$Ounces <- as.numeric(merged$Ounces)



colSums(is.na(merged))
```

# Now that we've organized the data provided we can get down to providing the information that
# you've requested.  In order to further understand MABco's presence in each state we have
# sorted the merged data by city and state.

```{r, echo=TRUE}
beer <- merged[order(merged$City),] 
```

# Given the merged results DBI was able to confirm that MABco's presence in each U.S.
# state is as follows:

```{r, echo=TRUE}
table(beer$State[beer$BreweryID])
```

# DBI was also tasked with providing a better understanding of MABco's beer offerings in
# terms of alcohol content (ABV).  We determined that the highest median ABV is 0.092, and
# is available in the states of Georgia and Florida when computing the average of all beer
# offerings in each state respectively.

# We have also provided median IBU values for each state for reference in the table below.
```{r, echo=TRUE}

##Create columns for both measurements to allow for quick lookups in the future on a per tuple basis
beer$StateABVMedian <- sapply(beer$ABV[beer$State], median)
beer$StateIBUMedian <- sapply(beer$IBU[beer$State], median)

##Create a list of every state included and store it in a variable
allStates <- unique(beer$State[beer$BreweryID])

##Use that list to index the new columns and create a data frame to display the findings
mediansByState <- data.frame(State=allStates, ABV_median = beer$StateABVMedian[allStates], IBU_median = beer$StateIBUMedian[allStates])

##Display the ne data frame that holds the median data
mediansByState
``` 

# To help MABco better understand these results we've plotted them graphically.
```{r, echo=TRUE}
library(ggplot2)
##Create a plot with the staes along the x-axis
p <- ggplot2::ggplot(mediansByState, aes(x = State))

##Create and add the y-axis. Also, give it a low alpha value and change it's fill color
p <- p + geom_bar(aes(y = IBU_median, color = "IBU"), stat="identity", alpha = .5 , fill="blue")

##Create and add the second y-axis. Also, give it a low alpha value and change it's fill color
##Note that the extra 100 multiplier is to turn the scale of those data into a percentage scale
p <- p + geom_bar(aes(y = ABV_median*100*6, color = "ABV"), alpha = .5 , fill="red", stat="identity")

##Create the right side second axis and give it the inverse scale used to transform the data
p <- p + scale_y_continuous(sec.axis = sec_axis(~./6, name = "ABV")) 

##Display the ggplot and marvel at its splendor
p
```


# Upon review of the results we derived from looking at the median alcohol content (ABV),
# as well as the median bitterness levels (IBU) we are able to confirm that the states with
# beer offerings with the highest ABV are Georgia and Florida (with median ABVs of 0.092) as
# previously confirmed.  

#However the beer with the highest ABV is Lee Hill Series Vol. 5 - 
# Belgian Style Quadrupel Ale (0.128 ABV).  This beer is sold in COLORADO.
```{r, echo=TRUE}
##Need code chunk here producing this result??##
```

# We are further able to confirm that the states with the most bitter beers as measured in
# IBU's are NY, OH, NC, OR, PA, NM, MT, SC, and NJ respectively, with each state returning
# a median IBU score of 135.  However, the individual state returning the highest IBU (138)
# is OREGON (Bitter Bitch Imperial IPA)
```{r, echo=TRUE}
##Need code chunk here producing this result??##
```

# Is this working as expected?
```{r, echo=TRUE}
beer$State[which.max(beer$ABV)]

beer$State[which.max(beer$IBU)]
```

##Questions/Whats left?? - This section should be deleted prior to finalization##
# 1) It looks like the code chunk above regarding NA's helps to alleviate our NA's, but we #didn't provide the
# initial total for NA's which was the ask.  Do we need to do that above?
# 2) I am not sure the ggplot above is performing as expected? Lets discuss.
# 3) Need to create a scatterplot to discuss the relationship between ABV and IBU and #summarize.



